Modal Abstractions for Virtualizing Memory Addresses

### 2024-07-24 examples.tex
figure 10

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry+KERNBASE}  }_{\rtv}$
$\specline{\texttt{rax} \mapsto_{\textsf{r}} \textsf{\_} \ast \lnot(\textsf{entry\_present entry\_val} }_{\rtv}$
$\specline{\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{entry entryv} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entry})}{\textsf{level}} }$
;;void pte_initialize(pte_t *entry) {
...
;;set up the stack    
;;allocate a full zeroed page for 512 8-byte entries 
;;pte_t *local = kalloc();
callq  81 <pte_initialize+0x11>	7d: R_X86_64_PLT32	kalloc-0x4 (*\label{line:call_to_kalloc}*)
mov    rax,-0x10[rbp] 
;;entry->pfn = PTE_ADDR_TO_PFN((uintptr_t) local);
mov    -0x10[rbp],rax
mov    -0x8[rbp],rdi
mov    rax,[rdi]
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry+KERNBASE} \ast \texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr}  }_{\rtv}$
$\specline{\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,}} \textsf{entry entryv} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entry})}{\textsf{level}} }$
$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \lnot(\textsf{entry\_present entry\_val}) }$
$\specline{ \textsf{entry+KERNBASE} \mapsto_{\mathsf{pte,1}} \mathsf{entry} \; \textsf{pfn\_set(entryv nextpaddr)}}$
$\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \begin{array}{l} \textsf{entry\_present(pte\_initialize(pfn\_set(entryv nextpaddr)))} -* \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511} \; \textsf{table\_root(pte\_initialized (pfn\_set (entryv nextpaddr))) + i * 8} \mapsto_{\textsf{id}} \textsf{level-1}    \end{array}  }_{\rtv}$
;;clean up the stack
...
\end{lstlisting}
\vspace{-1em}
\caption{Allocating a physical page }
\label{pteinitializespec}
\end{figure}
\todo[inline,color=blue]{Colin revision frontier}
Since we are using \textsf{pte\_initialize} for page-table address allocation, we must relate this newly allocated physical address to the identity mapping map $\Xi$ -- Specification Lines 34 and 38 in Figure \ref{fig:calltopteinitialize}. Focusing on the specification of \textsf{pte\_initialize} separately in Figure \ref{fig:pteinitializespec}, we right immediately realize that instead of seeing see a physical pointsto for the fresly page-table address (e.g. $\mathsf{nextpaddr} \mapsto_{\mathsf{p}} \mathsf{w64\_0}$) deliberately in the post-conditoin in Lines 15-16, we observe a full-ownership token representing the knowledge that a frame and all the entries indexed from this frame are freshly allocated with full-ownership to be a part of the identity map, $\Xi$. The soundness argument of this specification relies on the fact that these freshly allocated resources is part of an entry construction that has not been completed yet: the presence bit is set (Line 42 in Figure \ref{fig:calltopteinitialize}) after these freshly allocated resources are incorporated to the entry construction (Line 41 in Figure \ref{fig:calltopteinitialize}). In other words, the side condition, (*), acknowledges any access to the entry with these resources as \textit{invalid}, as the entry is revealed to shared-accesses when the presence bit is set.

@@

$\specline{ \textsf{entry}_{\textsf{pfn}}\textsf{+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val}  }_{\rtv}$
$\specline{  \textsf{entry}_{\textsf{pfn}} \textsf{+KERNBASE}  \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entrypnf\_val}  \textsf{entry}_{\textsf{pfn}}} \ast R_{\textsf{children}}$

>>>>

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry+KERNBASE}  }_{\rtv}$
$\specline{\texttt{rax} \mapsto_{\textsf{r}} \textsf{\_} \ast \lnot(\textsf{entry\_present entry\_val} }_{\rtv}$
$\specline{\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{entry entryv} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entry})}{\textsf{level}} }$
;;void pte_initialize(pte_t *entry) {
...
;;set up the stack    
;;allocate a full zeroed page for 512 8-byte entries 
;;pte_t *local = kalloc();
callq  81 <pte_initialize+0x11>	7d: R_X86_64_PLT32	kalloc-0x4 (*\label{line:call_to_kalloc}*)
mov    rax,-0x10[rbp] 
;;entry->pfn = PTE_ADDR_TO_PFN((uintptr_t) local);
mov    -0x10[rbp],rax
mov    -0x8[rbp],rdi
mov    rax,[rdi]
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry+KERNBASE} \ast \texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr}  }_{\rtv}$
$\specline{\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,}} \textsf{entry entryv} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entry})}{\textsf{level}} }$
$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \lnot(\textsf{entry\_present entry\_val}) }$
$\specline{ \textsf{entry+KERNBASE} \mapsto_{\mathsf{pte,1}} \mathsf{entry} \; \textsf{pfn\_set(entryv nextpaddr)}}$
$\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \begin{array}{l} \textsf{entry\_present(pte\_initialize(pfn\_set(entryv nextpaddr)))} -* \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511}} \; \textsf{table\_root(pte\_initialized (pfn\_set (entryv nextpaddr))) + i * 8} \mapsto_{\textsf{id}} \textsf{level-1}    \end{array}  }_{\rtv}$
;;clean up the stack
...
\end{lstlisting}
\vspace{-1em}
\caption{Allocating a physical page }
\label{pteinitializespec}
\end{figure}
\todo[inline,color=blue]{Colin revision frontier}
Since we are using \textsf{pte\_initialize} for page-table address allocation, we must relate this newly allocated physical address to the identity mapping map $\Xi$ -- Specification Lines 34 and 38 in Figure \ref{fig:calltopteinitialize}. Focusing on the specification of \textsf{pte\_initialize} separately in Figure \ref{fig:pteinitializespec}, we right immediately realize that instead of seeing see a physical pointsto for the fresly page-table address (e.g. $\mathsf{nextpaddr} \mapsto_{\mathsf{p}} \mathsf{w64\_0}$) deliberately in the post-conditoin in Lines 15-16, we observe a full-ownership token representing the knowledge that a frame and all the entries indexed from this frame are freshly allocated with full-ownership to be a part of the identity map, $\Xi$. The soundness argument of this specification relies on the fact that these freshly allocated resources is part of an entry construction that has not been completed yet: the presence bit is set (Line 42 in Figure \ref{fig:calltopteinitialize}) after these freshly allocated resources are incorporated to the entry construction (Line 41 in Figure \ref{fig:calltopteinitialize}). In other words, the side condition, (*), acknowledges any access to the entry with these resources as \textit{invalid}, as the entry is revealed to shared-accesses when the presence bit is set.

@@

$\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val}  }_{\rtv}$
$\specline{  \textsf{entry}_{\textsf{pfn}} \textsf{+KERNBASE}  \mapsto_{\textsf{vpte,qfrac}}   \textsf{entry}_{\textsf{pfn}} \; \textsf{entry\_pfn\_val} \ast R_{\textsf{children}}}_{\rtv}$
